### 2016-11-01 oldmort.Rnw
minor fix
run with fixed data

<<readtab>>=
load("data/vbTab.rda") 
load("data/vb.rda")
@

@@

@@

%The distribution over time periods.
<<getske6089>>=
vb6089 <- vbTab[as.numeric(as.character(vbTab$age)) >= 64.5, ]
vb6089$age <- factor(vb6089$age) # NOTE:: No change to '6589'!
vb6089$hisclass <- vb6089$hisclass60
@

@@

<<hazcalwold>>=
resw <- getPHglm(vb6089[vb6089$sex == "female", ],
              nper, labb)
resm <- getPHglm(vb6089[vb6089$sex == "male", ],
                nper, labb)
@

@@

Models are fitted separately for each ten-year time period, from  1961 to
2000. We stratify on HISCLASS
and use the covariates {\tt sex}, {\tt   civst}, and {\tt urban}
(Skellefteå town or outside). The cumulative
hazards at the end of the time period (60 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

@@

%The distribution over time periods.
<<41getske4060>>=
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, causes 19."
scap1w <-  "Cumulative hazards, ages 40--64, women, causes 19."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, causes 19."
scap1m <-  "Cumulative hazards, ages 40--64, men, causes 19."
##
cap12 <- "Total hazard of dying before age 60 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom), causes 19."
scap12 <- "Total hazard by HISCLASS, ages 40--64, causes 19."    
cap12log <- "Total hazard of dying before age 60 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, causes 19."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, causes 19."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, causes 19."
scap2w <-  "Cumulative hazards, ages 65--89, women, causes 19."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, causes 19."
scap2m <-  "Cumulative hazards, ages 65--89, men, causes 19."
##
cap22 <- "Total hazard of dying before age 90 for a 60 year old person by HISCLASS and decade, women (top) and men (bottom), causes 19."
scap22 <- "Total hazard by HISCLASS, ages 65--89, causes 19."    
cap22log <- "Total hazard of dying before age 90 for a 60 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, causes 19."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, causes 19."

@@

The statistical significance of the covariates are shown in Table~\ref{tab:liswom}
<<41liswom, results = 'asis'>>=
fitw1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "female", ])
##cat("Women:\n")
drw <- drop1(fitw1, test = "Chisq")
xtable(drw, caption = "Effect of covariates for women 40-64, 1990-2005, causes 19.", label = "tab:liswom")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

@@

<<41liswomfig, fig.cap = "Effect sizes, women 40-64, 1990-2005, causes 19.", fig.height = 4>>=
source("R/plotWEduc.R")
plotWEduc(fitw1)
@

@@

<<41lismen, results='asis'>>=
fitm1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "male", ])
##cat("\nMen:\n")
drm <- drop1(fitm1, test = "Chisq")
xtable(drm, caption = "Effect of covariates for men 40-64, 1990-2005, causes 19.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<41lismanfig, fig.cap = "Effect sizes, men 40-64, 1990-2005, causes 19.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm1)
@

<<41plotliss1, fig.cap = "Cumulative hazards, ages 40--64, 2002--2013, causes 19.", fig.height = 4>>=
#oldpar <- par(mfrow = c(1, 2))
#plot(fitw1, fn = "cum", col = dcol, main = "Women")
#plot(fitm1, fn = "cum", col = dcol, main = "Men")
#par(oldpar)
##summary(fitw1)
##summary(fitm1)
@

@@

%The distribution over time periods.
<<41getske6089>>=
vb6089 <- vbTab[as.numeric(as.character(vbTab$age)) >= 64.5, ]
vb6089$age <- factor(vb6089$age) # NOTE:: No change to '6589'!
@

@@

<<41hazcalwold>>=
resw <- getPHglm(vb6089[vb6089$sex == "female", ],
              nper, labb)
resm <- getPHglm(vb6089[vb6089$sex == "male", ],
                nper, labb)
@

@@

The statistical significance of the covariates are shown in Table~\ref{tab:liswom}
<<41liswom2, results = 'asis'>>=
fitw2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "female", ])
##cat("Women:\n")
drw2 <- drop1(fitw2, test = "Chisq")
xtable(drw2, caption = "Effect of covariates for women 65-89, 1990-2005, causes 19.", label = "tab:liswom")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

@@

<<41liswomfig2, fig.cap = "Effect sizes, women 65-89, 1990-2005, causes 19.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitw2)
@

@@

<<41lismen2, results='asis'>>=
fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "male", ])
##cat("\nMen:\n")
drm2 <- drop1(fitm2, test = "Chisq")
xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005, causes 19.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<41lismanfig2, fig.cap = "Effect sizes, men 65-89, 1990-2005, causes 19.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm2)
@

>>>>

<<readtab>>=
load("data/vbTab.rda") 
vbTab$period <- factor(vbTab$period)
load("data/vb.rda")
vb$period <- factor(vb$period)
@

@@

We will use as a summary measure the value of the cumulative hazards at
age 65.

@@

%The distribution over time periods.
<<getske6089>>=
vb6589 <- vbTab[as.numeric(as.character(vbTab$age)) >= 64.5, ]
vb6589$age <- factor(vb6589$age) # NOTE:: No change to '6589'!
vb6589$hisclass <- vb6589$hisclass60
@

@@

<<hazcalwold>>=
resw <- getPHglm(vb6589[vb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm(vb6589[vb6589$sex == "male", ],
                nper, labb)
@

@@

Models are fitted separately for each ten-year time period, from  1981 to
2013. We stratify on HISCLASS
and use the covariates {\tt sex}, {\tt   civst}, and {\tt urban}
(Skellefteå town or outside). The cumulative
hazards at the end of the time period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

@@

%The distribution over time periods.
<<41getske4060>>=
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cardiovascular death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cardiovascular death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cardiovascular death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cardiovascular death."
##
cap12 <- "Total hazard of dying before age 60 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cardiovascular death."    
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cardiovascular death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cardiovascular death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cardiovascular death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cardiovascular death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cardiovascular death."
##
cap22 <- "Total hazard of dying before age 90 for a 60 year old person by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap22 <- "Total hazard by HISCLASS, ages 65--89, cardiovascular death."    
cap22log <- "Total hazard of dying before age 90 for a 60 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cardiovascular death."

@@

The statistical significance of the covariates are shown in Table~\ref{tab:liswom}
<<41liswom, results = 'asis'>>=
fitw1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "female", ])
##cat("Women:\n")
drw <- drop1(fitw1, test = "Chisq")
xtable(drw, caption = "Effect of covariates for women 40-64, 1990-2005, cardiovascular death.", label = "tab:liswom")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

@@

<<41liswomfig, fig.cap = "Effect sizes, women 40-64, 1990-2005, cardiovascular death.", fig.height = 4>>=
source("R/plotWEduc.R")
plotWEduc(fitw1)
@

@@

<<41lismen, results='asis'>>=
fitm1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "male", ])
##cat("\nMen:\n")
drm <- drop1(fitm1, test = "Chisq")
xtable(drm, caption = "Effect of covariates for men 40-64, 1990-2005, cardiovascular death.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<41lismanfig, fig.cap = "Effect sizes, men 40-64, 1990-2005, cardiovascular death.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm1)
@

<<41plotliss1, fig.cap = "Cumulative hazards, ages 40--64, 2002--2013, cardiovascular death.", fig.height = 4>>=
#oldpar <- par(mfrow = c(1, 2))
#plot(fitw1, fn = "cum", col = dcol, main = "Women")
#plot(fitm1, fn = "cum", col = dcol, main = "Men")
#par(oldpar)
##summary(fitw1)
##summary(fitm1)
@

@@

%The distribution over time periods.
<<41getske6089>>=
vb6589 <- vbTab[as.numeric(as.character(vbTab$age)) >= 64.5, ]
vb6589$age <- factor(vb6589$age) # NOTE:: Now change to '6589'!
@

@@

<<41hazcalwold>>=
resw <- getPHglm(vb6589[vb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm(vb6589[vb6589$sex == "male", ],
                nper, labb)
@

@@

The statistical significance of the covariates are shown in Table~\ref{tab:liswom}
<<41liswom2, results = 'asis'>>=
fitw2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "female", ])
##cat("Women:\n")
drw2 <- drop1(fitw2, test = "Chisq")
xtable(drw2, caption = "Effect of covariates for women 65-89, 1990-2005, cardiovascular death.", label = "tab:liswom")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

@@

<<41liswomfig2, fig.cap = "Effect sizes, women 65-89, 1990-2005, cardiovascular death.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitw2)
@

@@

<<41lismen2, results='asis'>>=
fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "male", ])
##cat("\nMen:\n")
drm2 <- drop1(fitm2, test = "Chisq")
xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005, cardiovascular death.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<41lismanfig2, fig.cap = "Effect sizes, men 65-89, 1990-2005, cardiovascular death.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm2)
@
